*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-29 Code/Manuscript_O2.Rmd
more minor edits

Controlled growth experiments were performed using PSI Multicultivators (Figure \@ref(fig:LabeledMC); MCMIX-OD or MC1000-OD, Brno, Czech Republic).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (White LED, 660 nm, 450 nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~)
gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system (Qubit Systems Inc, Kingston, Ontario, Canada). [O~2~] in situ was verified using oxygen optodes (PyroScience,
Germany) inserted into tubes for
real-time measurements, with a temperature probe in
the aquarium of the bioreactor to correct [O~2~] measures for temperature
fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the
salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05%
CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

## Growth rate analysis {.unnumbered}
Data files (.csv) saved from the Multicultivator software were imported into R-Studio [@R-Studio] for data management [@R-tidyverse], growth rate calculations CITATION, comparisons of model fits (get Platt CITATION), and generation of
figures using `ggplot2` [@R-ggplot2]. The chlorophyll proxy
optical density (OD680 - OD720;  ΔOD) was used to determine the chlorophyll-specific exponential growth
rate (µ, d^-1^) for each treatment combination.  We first used a rolling mean from the RStudio `zoo` package [@R-zoo] to calculate the average ΔOD data over a 1-hour window to lower influence of outlier points and remove data points collected during post stationary phase, if applicable. We used the Levenberg-Marquardt algorithm
[@bellaviaLevenbergMarquardtMethod2018] modification of the non-linear
least squares,  using the R package `minpack.lm`
[@R-minpack.lm], to fit a logistic
equation \@ref(eq:GrowthRate)

@@

\begin{equation}
  Cumulative diel PUR = \frac{PUR × 3600 × Photoperiod}{2}
  (\#eq:SumPURDay)
\end{equation}

@@

MED4PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MED4 (High-Light (HLI) near surface clade) vs. photoperiod (h).  **",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

SS120PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 (Low-Light (LLIII) deep ocean clade) vs. photoperiod (h). **",
                     "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

MIT9313PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 (Low-Light (LLIV) deep ocean clade) vs. photoperiod (h). **",
                       "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

 BluevsRedPurFitsPlots_cap <- glue("**Chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4, SS120 and MIT9313.",
                    "Shapes show the imposed photoperiod (h); 4 h (solid square),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Harrison and Platt [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] 4 parameter model fit to 660 nm (red lines) and 450 nm (blue lines) growth data for each combination of strain and dissolved oxygen shown with solid lines (significantly different fits, pvalue < 0.05) or dashed lines (not significantly different fits, pvalue > 0.05) using one-way ANOVA.",
                       .sep = " ")

@@

AllStrainPurFitsPlots_cap <- glue("**Chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4, SS120 and MIT9313.",
                    "Shapes show the imposed photoperiod (h); 4 h (solid square),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; white LED (black symbols), 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Harrison and Platt [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] 4 parameter model fit to data pooled for each combination of strain and dissolved oxygen shown with solid lines. Separate models fit to photoperiod data and shown if significantly different from the pooled model using one-way ANOVA; 4 h (long dashed line); 8 h (dotted line); 12 h (dashed line); and 16 h (dot dashed line).",
                       .sep = " ")

>>>>

Controlled growth experiments were performed using MCMIX-OD or MC1000-OD PSI Multicultivators (Figure \@ref(fig:LabeledMC); PSI, Drásov, Czech Republic).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (White LED, 660 nm, 450 nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~)
gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system (Qubit Systems Inc, Kingston, Ontario, Canada). [O~2~] in situ was verified using oxygen optodes (PyroScience,
Germany) inserted into tubes for
real-time measurements, with a temperature probe in
the aquarium of the bioreactor to correct [O~2~] measures for temperature
fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the
salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05%
CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

## Growth rate analysis {.unnumbered}
Data files (.csv) saved from the Multicultivator software were imported into R-Studio [@R-Studio] for data management [@R-tidyverse], growth rate calculations CITATION, comparisons of model fits [@harrisonPhotosynthesisirradianceRelationshipsPolar1986], and generation of
figures using `ggplot2` [@R-ggplot2]. The chlorophyll proxy
optical density (OD680 - OD720;  ΔOD) was used to determine the chlorophyll specific growth
rate (µ, d^-1^) for each treatment combination.  We first used a rolling mean from the RStudio `zoo` package [@R-zoo] to calculate the average ΔOD data over a 1-hour window to lower influence of outlier points and remove data points collected during post stationary phase, if applicable. We used the Levenberg-Marquardt algorithm
[@bellaviaLevenbergMarquardtMethod2018] modification of the non-linear
least squares,  using the R package `minpack.lm`
[@R-minpack.lm], to fit a logistic
equation \@ref(eq:GrowthRate)

@@

\begin{equation}
  Cumulative~diel~PUR = \frac{PUR × 3600 × Photoperiod}{2}
  (\#eq:SumPURDay)
\end{equation}

@@

MED4PAR_cap <- glue("**chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MED4 (High-Light (HLI) near surface clade) vs. photoperiod (h).  **",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

SS120PAR_cap <- glue("**chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 (Low-Light (LLIII) deep ocean clade) vs. photoperiod (h). **",
                     "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

MIT9313PAR_cap <- glue("**chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 (Low-Light (LLIV) deep ocean clade) vs. photoperiod (h). **",
                       "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

 BluevsRedPurFitsPlots_cap <- glue("**chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4, SS120 and MIT9313.",
                    "Shapes show the imposed photoperiod (h); 4 h (solid square),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Harrison and Platt [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] 4 parameter model fit to 660 nm (red lines) and 450 nm (blue lines) growth data for each combination of strain and dissolved oxygen shown with solid lines (significantly different fits, pvalue < 0.05) or dashed lines (not significantly different fits, pvalue > 0.05) using one-way ANOVA.",
                       .sep = " ")

@@

AllStrainPurFitsPlots_cap <- glue("**chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4, SS120 and MIT9313.",
                    "Shapes show the imposed photoperiod (h); 4 h (solid square),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; white LED (black symbols), 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Harrison and Platt [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] 4 parameter model fit to data pooled for each combination of strain and dissolved oxygen shown with solid lines. Separate models fit to photoperiod data and shown if significantly different from the pooled model using one-way ANOVA; 4 h (long dashed line); 8 h (dotted line); 12 h (dashed line); and 16 h (dot dashed line).",
                       .sep = " ")
